---
title: "Introductory Statistics for Economics"
author: "Duong Trinh"
date: "Feb 2022"
output:
  beamer_presentation:
    keep_tex: yes
    slide_level: 2
  slidy_presentation: default
  ioslides_presentation: default
subtitle: 'ECON1013: LAB 2'
fontsize: 10pt
institute: University of Glasgow
---

## Picture the Scenario

```{=tex}
\begin{itemize}
    \item \textbf{Objective:} 
    \begin{itemize}
        \item[$\square$] illustrate how the statistic $\bar{X}_n$ varies with repeated samples
        \item[$\square$]  describe what might happen if we repeat the entire sampling process  and  compute the sample average again and again.
        \item[$\square$] illustrate the CLT and the LLN
    \end{itemize}
    \vspace{3mm}
    \item \textbf{Implementation:} 
    \begin{itemize}
        \item[$\square$] We use a \textit{pseudo-random number generator}, already built in R, to simulate the samples. Simulations allow the user to perform experiments and answer questions in a rapid manner.
    \end{itemize}
\end{itemize}
```
# PRELIMINARIES

## The normal distribution

For each distribution, R provides functions that allow users calculate value of: quantile, probability and generate random variable of the distribution.

```{=tex}
\begin{itemize}
    \item To calculate the \textit{probability} of $P(X \leq 0.05)$ where $X \sim N(0,1)$: 
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
pnorm(q = 0.05, mean = 0, sd = 1) 
```

```{=tex}
\begin{itemize}
    \item To calculate the \textit{probability} of $P(X \leq 1)$ where $X \sim N(4,3^2)$:
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
pnorm(q = 1, mean = 4, sd = 3) 
```

## The normal distribution

For each distribution, R provides functions that allow users calculate value of: quantile, probability and generate random variable of the distribution.

```{=tex}
\begin{itemize}
    \item To calculate the \textit{quantile} at $p = 0.05$ of $N(0,1)$:
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
qnorm(p = 0.05,mean = 0,sd = 1) 
```

```{=tex}
\begin{itemize}
    \item To calculate the \textit{quantile} at $p = 0.05$ of $N(4,3^2)$:
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
qnorm(p = 0.05,mean = 4,sd = 3) 
```

## The normal distribution

For each distribution, R provides functions that allow users calculate value of: quantile, probability and generate random variable of the distribution.

```{=tex}
\begin{itemize}
    \item To calculate value of \textit{density function} of $N(0,1)$ at $0.4$:
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
dnorm(x = 0.4, mean = 0, sd = 1) 
```

```{=tex}
\begin{itemize}
    \item To calculate value of \textit{density function} of $N(4,3^2)$ at $0.4$:
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
dnorm(x = 0.4, mean = 4, sd = 3)    
```

## The normal distribution

For each distribution, R provides functions that allow users calculate value of: quantile, probability and generate random variable of the distribution.

```{=tex}
\begin{itemize}
    \item To generate randomly $1000$ numbers from $N(0,1)$:
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
R <- rnorm(n = 1000, mean = 0, sd = 1)
head(R)
length(R)
summary(R)
```

## Loops (\texttt{for}) and conditional expressions (\texttt{if})

```{=tex}
\begin{itemize} 
    \item Create the vector GENDER where the entries are either \texttt{male} or \texttt{female}
\end{itemize}
```
```{r eval=TRUE,message=FALSE}
GENDER = c('male', 'male', 'female', 'male', 'female')
```

```{=tex}
\begin{itemize} 
    \item Define the random variable $X$ taking value 1 if the individual is a male, and 0 if a female.
\end{itemize}
```
```{r eval=TRUE,message=FALSE,results=FALSE}
X=numeric(0)
for (i in 1:length(GENDER)) {
  if (GENDER[i] == 'male') {
    X[i] = 1
  } else {
    X[i]=0
  }
}
```

# Exercise 1: The Central Limit Theorem

## Exercise 1: The Central Limit Theorem

```{=tex}
\begin{itemize}
  \item Simulate 10,000 samples of size $n$ from an exponential distribution with mean $\mu=10$. Consider different sample sizes, e.g. $n = 10, 30, 500$. 
  \item For each sample, compute the standardized sample average: $$
  \tilde{X}=\frac{\bar{X}-\mathbb{E}(\bar{X})}{\sigma_{\bar{x}}}
$$

  \item Compute the empirical quantiles  of the sample distribution of $\tilde{X}$. Compare the results with the theoretical quantiles of the standard normal random variable $Z$.
  \item Compute quantile $z$ satisfying $P(Z<z)=0.975$. Calculate the percentage of values of $\tilde{X}$ greater than $z$.
\end{itemize}
```
Comment on your results.

## Exercise 1: The Central Limit Theorem

Specify the parameters of the experiment

```{r}
# "n" is the sample size. Try n = 10, 30, 500
n = 10
# "R" is the number of replications
R = 10000
# Parameter exponential
mu = 10 # mean
```

To reproduce results, setting a seed in R means to initialize a pseudorandom number generator

```{r}
set.seed(12) # take an (arbitrary) integer 
```

## Exercise 1: The Central Limit Theorem

Simulate $R$ samples of size $n$ from an exponential distribution with mean $\mu = 10$ and compute the standardized sample average for each sample:

$$
Y=\frac{\bar{X}-\mathbb{E}(\bar{X})}{\sigma_{\bar{x}}}
$$

```{r}
# Create vector Y (pre-allocation) 
Y = numeric(0) 
# Loop through R replications
for (i in 1:R) {
  X = rexp(n,1/mu) # The sample will change for every "i"
  # Compute standardized sample average
  Y[i] = (mean(X)-mu)/(mu/sqrt(n))
}
```

## Exercise 1: The Central Limit Theorem

To assess the how well the sampling distribution of X is approximated by the standard normal we compare the quantile of of X and the quantile for a standard normal distribution.

## Exercise 1: The Central Limit Theorem

```{r}
# Empirical quantiles
quantile(Y)
```

```{r}
# Population quantiles of a standard normal distribution
qnorm(p = 0.25, mean = 0, sd = 1)
qnorm(p = 0.5, mean = 0, sd = 1)
qnorm(p = 0.75, mean = 0, sd = 1)

```

## Exercise 1: The Central Limit Theorem

```{r}
hist(Y, prob = T)
```

## Exercise 1: The Central Limit Theorem

```{r}
qqnorm(Y,main='Normal(0,1)');qqline(Y) #normal plots
```

## Exercise 1: The Central Limit Theorem

```{r}
# Compute the quantile z satisfying P(Z<z) = 0.975
z = qnorm(p = 0.975, mean = 0, sd = 1)
z
# The percentage of values of Y greater than z
sum(Y>z)/R*100 
```

If the sample is well approximated by a standard normal distribution, the realization of $Y$ should be greater than $z$ : $(1-0.975*R) = 250$ times, or $2.5\%$ of the times.

## Exercise 1: The Central Limit Theorem

```{r eval=TRUE,message=FALSE,warning=FALSE,echo=FALSE}
library(ggplot2)
library(gridExtra)
set.seed(12)
# "R" is the number of replications
R = 10000
# "n" is the sample size. Try n=10,30,500
for (n in c(10,30,500)) {

  # Create the vector Y (pre-allocation) to store the R standardized sample averages
Y = numeric(0) 
for (i in 1:R) {
  mu = 10
  X = rexp(n,1/mu) # The sample will change for every "i"
  # Compute standardized sample average
  Y[i] = (mean(X)-mu)/(mu/sqrt(n))
}

Y <- data.frame(Y)

if (n == 10) {
h1 <- ggplot(Y,aes(Y))+
  geom_histogram(color = "blue",fill="white")+
  geom_vline(aes(xintercept = mean(Y)), color = "red")+
  ggtitle("N = 10")

qq1 <- ggplot(Y, aes(sample = Y))+
  stat_qq(color = 'blue') + stat_qq_line(color = 'red')+
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles")+
  ggtitle("N = 10")
}
else if (n == 30) {
  h2 <- ggplot(Y,aes(Y))+
    geom_histogram(color = "blue",fill="white")+
    geom_vline(aes(xintercept = mean(Y)), color = "red")+
    ggtitle("N = 30")
  
  qq2 <- ggplot(Y, aes(sample = Y))+
    stat_qq(color = 'blue') + stat_qq_line(color = 'red')+
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles")+
    ggtitle("N = 30")
}
else {
  h3 <- ggplot(Y,aes(Y))+
    geom_histogram(color = "blue",fill="white")+
    geom_vline(aes(xintercept = mean(Y)), color = "red")+
    ggtitle("N = 500")
  
  qq3 <- ggplot(Y, aes(sample = Y))+
    stat_qq(color = 'blue') + stat_qq_line(color = 'red')+
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles")+
    ggtitle("N = 500")
}
}
grid.arrange(h1,h2,h3,qq1,qq2,qq3,ncol=3,nrow=2)
```

## The Central Limit Theorem - Reflection

\textcolor{blue}{The distribution of the sample average $\bar{X_n}$ will be approximately normal as long as the sample size $n$ is large enough.}\

# Exercise 2: The Law of Large Numbers

## Exercise 2: The Law of Large Numbers

```{=tex}
\begin{itemize}
  \item Compute the sample averages 
  $$
  \bar{x}_n = \frac{\sum_{i=1}^n x_i}{n}
  $$
for $n$ draws from the exponential distribution considered in Exercise 1, for $n=1,2,\dots,1000$.  
  \item Plot the sample averages. What do you conclude about the variability of the sample average?
\end{itemize}
```
## Exercise 2: The Law of Large Numbers

```{r eval=TRUE,message=FALSE,fig.show='hide'}
set.seed(12) # to reproduce results
N = 1000 # maximum sample size
mu = 10 # parameter exponential 
Y = numeric(0) # pre-allocation 
for (i in 1:N) {
  X = rexp(i,1/mu) # 'i' draws from Exp(1/mu)
  Y[i] = mean(X) # sample average
}
plot(Y, xlab = "Sample size n", ylab = "Sample average Y")
lines(rep(mu,N), col = 'red')
```

## Exercise 2: The Law of Large Numbers

```{r eval=TRUE,message=FALSE,echo=FALSE}
set.seed(12) # to reproduce results
N = 1000 # maximum sample size
mu = 10 # parameter exponential 
Y = numeric(0) # pre-allocation 
for (i in 1:N) {
  X = rexp(i,1/mu)
  Y[i] = mean(X)
}
plot(Y, xlab = "Sample size n", ylab = "Sample average Y")
lines(rep(mu,N), col = 'red')
```

## Exercise 2: The Law of Large Numbers

```{r eval=TRUE,message=FALSE,warning=FALSE,echo=FALSE}
library(ggplot2)
set.seed(12)
mu = 10

N = 100
Y1 = numeric(0)
  for (i in 1:N) {
    X = rexp(i,1/mu)
    Y1[i] = mean(X)
  }
n1 = 1:N
Yn1 <- data.frame(Y1,n1)

p1 <- ggplot(Yn1,aes(n1,Y1))+
  geom_point(shape = 2, color = 4, size = 1)+
  geom_hline(yintercept = mu, color = 'red')+
  labs(x = "Sample size n", y = "Sample average Y")+
  ggtitle("N = 100")+
  coord_cartesian(ylim = c(0, 26))

N = 1000
Y2 = numeric(0)
for (i in 1:N) {
  X = rexp(i,1/mu)
  Y2[i] = mean(X)
}
n2 = 1:N
Yn2 <- data.frame(Y2,n2)

p2 <- ggplot(Yn2, aes(n2,Y2))+
  geom_point(shape = 2, color = 4, size = 1)+
  geom_hline(yintercept = mu, color = 'red')+
  labs(x = "Sample size n", y = "Sample average Y")+
  ggtitle("N = 1000")+
  coord_cartesian(ylim = c(0, 26))

N = 10000
Y3 = numeric(0)
for (i in 1:N) {
  X = rexp(i,1/mu)
  Y3[i] = mean(X)
}
n3 = 1:N
Yn3 <- data.frame(Y3,n3)

p3 <- ggplot(Yn3,aes(n3,Y3))+
  geom_point(shape = 2, color = 4, size = 1)+
  geom_hline(yintercept = mu, color = 'red')+
  labs(x = "Sample size n", y = "Sample average Y")+
  ggtitle("N = 10000")+
  coord_cartesian(ylim = c(0, 26))

library(gridExtra)
grid.arrange(p1,p2,p3,ncol=3,nrow=1)
```

## The Law of Large Numbers - Reflection

\textcolor{blue}{Given a random sample of size $n$ from a population, the sample average $\bar{X_n}$ will approach the population mean $\mu$ as the sample size $n$ becomes large.}
